---
title: "Anomalous diffusion of monomers after two DSBs"
author: "Ignacio Madrid"
date: "June 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include = FALSE}
library(ggplot2)
library(reshape2)
```

## Anomalous diffusion keeping and removing CLs in the DF

```{r data, include = FALSE}
res_keep <- read.csv("../DNA_Religation/projects/AnomalousExponentAfterDSBs/results/2018_06_22_17_39_trackAlphaAfterDSB_vsNc.csv")
res_remo <- read.csv("../DNA_Religation/projects/AnomalousExponentAfterDSBs/results/2018_06_28_12_42_trackAlphaAfterDSB_vsNc.csv")
```

```{r extract_alphas, include = FALSE}
get_alpha <- function(res){
alphas <- res[, grepl("alpha.", names(res))]
alphas$nc <- res$Nc
alphas$alpha_monomers <- NULL
return(alphas)
}

get_ends.alpha <- function(res){
alphas <- res[, c("alpha.30","alpha.31","alpha.68","alpha.69")]
alphas$nc <- res$Nc
return(alphas)
}

alphas.k <- get_alpha(res_keep)
alphas.r <- get_alpha(res_remo)
```

```{r alphas_dataframes, include = FALSE}
alphas.k.melted <- melt(alphas.k, id.vars = "nc", variable.name = "monomer", value.name = "alpha")
alphas.r.melted <- melt(alphas.r, id.vars = "nc", variable.name = "monomer", value.name = "alpha")

alphas.k.melted$monomer <- c(sapply(seq(100),FUN = function(x){rep(x,29)}))
alphas.r.melted$monomer <- c(sapply(seq(100),FUN = function(x){rep(x,29)}))

alphas.avg <- data.frame(monomer = seq(100),
                         keep = sapply(alphas.k[,!names(alphas.k) %in% ("nc")], mean),
                         remo = sapply(alphas.r[,!names(alphas.r) %in% ("nc")], mean))

alphas.avg.melted <- melt(alphas.avg, id.vars = c("monomer"), value.name = "alpha")

a1 = 31
a2 = 32
b1 = 69 
b2 = 70
```


The anomalous diffusion exponent $\alpha$ was extraited from the trajectories of each one of the 100 monomers of the polymer after the induction of two fixed DSBs at $A_1 = 46$ and $B_1 = 54$ (the first monomers is numerated $1$). The break is followed by the removal or not of the CLs in the DF, and finally a waiting relaxation time of 10000 steps (50 seconds) before start tracking the monomers' position. The same is performed for different values of $N_c$ between 2 and 58. $N_d = 3$ CLs were imposed in the DF to compare both scenarios.

Below we see $\alpha$ for each monomer when CLs are kept after the DSB.

```{r, echo = F}
ggplot(data = alphas.k.melted, aes(x = monomer, y = alpha, color = nc)) + 
  geom_point() + geom_line(aes(group = nc)) +
  geom_vline(xintercept = c(a1,a2,b1,b2)) +
  scale_color_gradient(low = "yellow", high = "red")
```

And when CLs are removed:

```{r, echo = F}
ggplot(data = alphas.r.melted, aes(x = monomer, y = alpha, color = nc)) + 
  geom_point() + geom_line(aes(group = nc)) +
  geom_vline(xintercept = c(a1,a2,b1,b2)) +
  scale_color_gradient(low = "yellow", high = "red")
```

And finally, averageing over all realisations of $N_c$ (30 CLs in average):

```{r, echo = F}
ggplot(data = alphas.avg.melted, aes(x= monomer, y = alpha, color = variable, shape = variable)) + 
  geom_point(size = 2) + geom_line(aes(group = variable), size = 1) +
  geom_vline(xintercept = c(a1,a2,b1,b2), linetype = 1, size = 1)
```


## $\alpha$ of DF ends as function of $N_c$

### Keeping CLs

Anomalous exponents of the ends:

```{r, echo = F}
alphas.k.melted$monomer <- as.factor(alphas.k.melted$monomer)
ggplot(data = alphas.k.melted[alphas.k.melted$monomer %in% c(a1,a2,b1,b2),], aes(x = nc, y = alpha, color = monomer)) + geom_point() + geom_line(aes(group = monomer))
```

Anomalous exponents of all the monomers

```{r, echo = F}
ggplot(data = alphas.k.melted, aes(x = nc, y = alpha, color = monomer)) + 
  geom_point() + geom_line(aes(group = monomer))
```


### Removing CLs

Anomalous exponents of the ends:

```{r, echo = F}
alphas.r.melted$monomer <- as.factor(alphas.r.melted$monomer)
ggplot(data = alphas.r.melted[alphas.r.melted$monomer %in% c(a1,a2,b1,b2),], aes(x = nc, y = alpha, color = monomer)) + geom_point() + geom_line(aes(group = monomer))
```

Anomalous exponents of all the monomers

```{r, echo = F}
ggplot(data = alphas.r.melted, aes(x = nc, y = alpha, color = monomer)) + 
  geom_point() + geom_line(aes(group = monomer))
```


<!-- ## Including the effect of selective volume exclusion ($\sigma = 0.1 \mu$m) -->

<!-- We'll consider now the same experiment but adding a selective exclusion sphere of radius $\sigma = 0.1 \mu$m around each one of the 4 DF ends. -->

```{r data ve, include = FALSE}
res_keep01 <- read.csv("../DNA_Religation/projects/AnomalousExponentAfterDSBs/results/2018_06_23_03_49_trackAlphaAfterDSB_vsNc.csv")
res_remo01 <- read.csv("../DNA_Religation/projects/AnomalousExponentAfterDSBs/results/2018_06_24_15_18_trackAlphaAfterDSB_vsNc.csv")

alphas.k.ve01 <- get_alpha(res_keep01)
alphas.r.ve01 <- get_alpha(res_remo01)

alphas.k.melted.ve01 <- melt(alphas.k.ve01, id.vars = "nc", variable.name = "monomer", value.name = "alpha")
alphas.r.melted.ve01 <- melt(alphas.r.ve01, id.vars = "nc", variable.name = "monomer", value.name = "alpha")

alphas.k.melted.ve01$monomer <- c(sapply(seq(100),FUN = function(x){rep(x,29)}))
alphas.r.melted.ve01$monomer <- c(sapply(seq(100),FUN = function(x){rep(x,29)}))

alphas.avg.ve01 <- data.frame(monomer = seq(100),
                         keep = sapply(alphas.k.ve01[,!names(alphas.k) %in% ("nc")], mean),
                         remo = sapply(alphas.r.ve01[,!names(alphas.r) %in% ("nc")], mean))

alphas.avg.melted.ve01 <- melt(alphas.avg.ve01, id.vars = c("monomer"), value.name = "alpha")
```

<!-- Below we see $\alpha$ for each monomer when CLs are kept after the DSB. -->

<!-- ```{r, echo = F} -->
<!-- ggplot(data = alphas.k.melted.ve01, aes(x = monomer, y = alpha, color = nc)) +  -->
<!--   geom_point() +  -->
<!--   geom_vline(xintercept = c(31,32,69,70)) + -->
<!--   scale_color_gradient(low = "yellow", high = "red") -->
<!-- ``` -->

<!-- And when CLs are removed: -->

<!-- ```{r, echo = F} -->
<!-- ggplot(data = alphas.r.melted.ve01, aes(x = monomer, y = alpha, color = nc)) +  -->
<!--   geom_point() +  -->
<!--   geom_vline(xintercept = c(31,32,69,70)) + -->
<!--   scale_color_gradient(low = "yellow", high = "red") -->
<!-- ``` -->

<!-- And finally, averageing over all realisations of $N_c$ (25 CLs in average): -->

<!-- ```{r, echo = F} -->
<!-- ggplot(data = alphas.avg.melted.ve01, aes(x= monomer, y = alpha, color = variable)) +  -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept = c(31,32,69,70)) -->
<!-- ``` -->

## Comparaison keeping/removing CLs $\times$ with/without VE (average over all $N_c$'s)

We compare the average curves for the four experimental scenarios:
```{r, echo = F}
alphas.avg.all <- data.frame(monomer = seq(100),
                             keep = alphas.avg$keep,
                             remo = alphas.avg$remo,
                             sigma0.1 = alphas.avg.ve01$keep)

alphas.avg.all.melted <- melt(alphas.avg.all, id.vars = c("monomer"), value.name = "alpha")

ggplot(data = alphas.avg.all.melted, aes(x= monomer, y = alpha, color = variable)) + 
  geom_line(size = 1) +
  geom_vline(xintercept = c(a1,a2,b1,b2))
```


```{r, echo = F}
ends.k <- melt(get_ends.alpha(res_keep), id.vars = "nc", value.name = "alpha", variable.name = "monomer")
ends.r <- melt(get_ends.alpha(res_remo), id.vars = "nc", value.name = "alpha", variable.name = "monomer")
ends.ve <- melt(get_ends.alpha(res_keep01), id.vars = "nc", value.name = "alpha", variable.name = "monomer")
ends.k$scenario <- "Keeping CLs"
ends.r$scenario <- "Removing CLs"
ends.ve$scenario <- "VE 0.1 (keeping CLs)"
ends.k$monomer <- ends.r$monomer <- ends.ve$monomer <- as.factor(c(sapply(c(a1,a2,b1,b2),FUN = function(x){rep(x,29)})))
ends <- rbind(ends.k, ends.r, ends.ve)

p <- ggplot(data = ends, mapping = aes(x = nc, y = alpha, color = monomer)) + 
  facet_grid(.~scenario, scale="free") + 
  geom_line(data = ends.k, stat = "identity", size = 1) + 
  geom_line(data = ends.r, stat = "identity", size = 1) + 
  geom_line(data = ends.ve, stat = "identity", size = 1)
p


```

<!-- ## Interactive comparaison (for each Nc) -->

<!-- ```{r, echo = F} -->
<!-- library(shiny) -->

<!-- ui <- fluidPage( -->
<!--  selectInput(inputId = "experiment",  -->
<!--              label = "Select experimental scenario",  -->
<!--              choices = c("Keep CLs in DF", "Remove CLs in DF")), -->
<!--  sliderInput(inputId = "nc",  -->
<!--              label = "Select Nc",  -->
<!--              min = 2, max = 48, value = 20,  -->
<!--              step = 2), -->
<!--  plotOutput(outputId = "plot")  -->
<!-- ) -->

<!-- server <- function(input, output) { -->
<!--  output$plot <- renderPlot({ -->
<!--  ################## -->
<!--    if (input$experiment == "Keep CLs in DF") { -->
<!--      df <- data.frame(monomer = seq(100), -->
<!--                       withoutVE = alphas.k.melted[alphas.k.melted$nc==input$nc,]$alpha, -->
<!--                       withVE.01 = alphas.k.melted.ve01[alphas.k.melted$nc==input$nc,]$alpha) -->
<!--    } -->
<!--    else{ -->
<!--      df <- data.frame(monomer = seq(100), -->
<!--                       withoutVE = alphas.r.melted[alphas.k.melted$nc==input$nc,]$alpha, -->
<!--                       withVE.01 = alphas.r.melted.ve01[alphas.k.melted$nc==input$nc,]$alpha) -->
<!--    } -->

<!--   df.melted <- melt(df, id.vars = "monomer", value.name = "alpha", variable.name = "VE") -->

<!--   ggplot(data = df.melted, aes(x = monomer, y = alpha, color = VE)) +  -->
<!--   geom_line() +  -->
<!--   geom_vline(xintercept = c(31,32,69,70))  -->
<!--  ################## -->
<!--  }) -->
<!-- } -->

<!-- shinyApp(ui = ui, server = server) -->
<!-- ``` -->